Syntax-Based Concept Alignment for Domain-Specific Machine Translation

### 2021-05-10 paper/paper.tex
oops the problem was a missing style file
started working on bibliography

\section{Introduction}
Grammar-based translation pipelines such as those based on Grammatical Framework (GF) \cite{dummy} have been successfully employed in domain-specific Machine Translation (MT).  
What makes these systems well suited to the task is the fact that, when we constrain ourselves to a specific domain, where precision is most often more important than coverage, they can provide strong guarantees in terms of grammatical correctness.

Nevertheless, lexical exactness is, in this context, just as important as grammaticality. 
An important part of the design of the Controlled Natural Language (CNL) the grammar in such a system describes becomes, then, the creation of a translation lexicon. 
In many cases, this is done for the most part manually, resulting in a time consuming task requiring significant linguistic knowledge. 
When the grammar is designed based on a corpus of parallel example sentences, it is possible to automate part of this process by means of statistical word and phrase alignment techniques \cite{dummy}. 
None of them is, however, suitable for the common case in which only a limited number of example data is available.

@@

As stated in the Introduction, most existing extraction solutions are based on statistical approaches and are, as a consequence, unsuitable for small datasets. 
Grammar-based approaches, making use of parallel treebanks and collectively referred to as \textit{tree-to-tree alignment methods}, have also been proposed \cite{dummy}, but have historically suffered from the inconsistencies between the various formalisms used to define the grammars for different languages and from the insufficient degree of robustness of existing parses.

This work is a new attempt in the same direction, enabled by two multilinguality-oriented grammar formalisms developed over the course of the last 25 years: Universal Dependencies (UD) \cite{dummy} and Grammatical Framework (GF) \cite{dummy}.

While both formalisms, the former a dependency and the latter a constituency grammar\todo{too quick, say more about both}, independently solve the former issue, UD is especially appealing since dependency trees are an easier target for parsing. 
As a consequence, several robust dependency parsers, such as \cite{dummy} and \cite{dummy} are available.

@@

% TODO: see why it does not compile
% Entries for the entire Anthology, followed by custom entries
\bibliography{custom}  
\bibliographystyle{plain}

>>>>

\section{Introduction}
Grammar-based translation pipelines such as those based on Grammatical Framework (GF) have been successfully employed in domain-specific Machine Translation (MT) \cite{scaling_up_gf}.  
What makes these systems well suited to the task is the fact that, when we constrain ourselves to a specific domain, where precision is most often more important than coverage, they can provide strong guarantees in terms of grammatical correctness.

Nevertheless, lexical exactness is, in this context, just as important as grammaticality. 
An important part of the design of the Controlled Natural Language (CNL) the grammar in such a system describes becomes, then, the creation of a translation lexicon. 
In many cases, this is done for the most part manually, resulting in a time consuming task requiring significant linguistic knowledge. 
When the grammar is designed based on a corpus of parallel example sentences, it is possible to automate part of this process by means of statistical word and phrase alignment techniques \cite{brown-etal-1993-mathematics, och-ney-2000-improved, dyer-etal-2013-simple}. 
None of them is, however, suitable for the common case in which only a limited number of example data is available.

@@

As stated in the Introduction, most existing extraction solutions are based on statistical approaches and are, as a consequence, unsuitable for small datasets. 
Grammar-based approaches, making use of parallel treebanks and collectively referred to as \textit{tree-to-tree alignment methods}, have also been proposed \cite{tiedemann2011bitext}, but have historically suffered from the inconsistencies between the various formalisms used to define the grammars for different languages and from the insufficient degree of robustness of existing parses.

This work is a new attempt in the same direction, enabled by two multilinguality-oriented grammar formalisms developed over the course of the last 25 years: Universal Dependencies (UD) \cite{ws-2019-universal} and Grammatical Framework (GF) \cite{scaling_up_gf}.

While both formalisms, the former a dependency and the latter a constituency grammar\todo{too quick, say more about both}, independently solve the former issue, UD is especially appealing since dependency trees are an easier target for parsing. 
As a consequence, several robust dependency parsers, such as \cite{straka-etal-2016-udpipe} and \cite{chen-manning-2014-fast} are available.

@@

% TODO: see why it does not compile
% Entries for the entire Anthology, followed by custom entries
\bibliography{anthology,custom}  
\bibliographystyle{acl_natbib}
